---
title: "Coconino County Health and Human Services Vital Record and Medical Examiner Data for Death by Suicide"
subtitle: "Supplemental Data"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  beamer_presentation:
    colortheme: "seahorse"
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  fig.width = 11
)
```

```{r libraries}
library(here)
library(tidyverse)
library(pins)
library(knitr)
library(RColorBrewer)
library(scales)
library(cowplot)
```

```{r theme}
theme_suicide_data_report <- theme_classic() +
  theme(
    axis.line = element_line(size = .5, color = "#979797"),
    axis.ticks = element_line(size = .5, color = "#979797"),
    aspect.ratio = 3 / 5.5,
    text = element_text(
      size = 14,
      color = "#3b3b3b"
    )
  )
```

```{r read}
# load pin board
suicide_data <- board_folder("S:/HIPAA Compliance/SAS Files/Coconino Deaths/Suicide/data-raw")

# read complete data set
azdhs_suicide_data_extract <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract")

# lower limit
analysis_year_range_ll <- as.character(min(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# upper limit
analysis_year_range_ul <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# current year
analysis_year_current <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_current)))

# years of analysis
analysis_year_range <- c(as.character(analysis_year_range_ll:analysis_year_range_ul))

# create a character string showing range of years included in report
years_of_analysis <- str_c(analysis_year_range_ll,
  analysis_year_range_ul,
  sep = " - "
)

# read geospatial rate by zip code
gis_rate_zipcode_spatial <- suicide_data %>%
  pin_read("gis_rate_zipcode_spatial")

# read geospatial coconino county boundary
gis_coco_county_spatial <- suicide_data %>%
  pin_read("gis_coco_county_spatial")

# read geospatial coconino county primary roads
gis_coco_county_roads_spatial <- suicide_data %>%
  pin_read("gis_coco_county_roads_spatial")

# read geocoded injury locations 
geocoded_injury_location_spatial <- suicide_data %>%
  pin_read("geocoded_injury_location_spatial")

# read all cause mortality 
azdhs_all_cause_mortality_coconino_resident_count <- suicide_data %>%
  pin_read("azdhs_all_cause_mortality_coconino_resident_count")

# read all cause mortality military denominator
azdhs_all_cause_mortality_coconino_military_count <- suicide_data %>%
  pin_read("azdhs_all_cause_mortality_coconino_military_count")
azdhs_all_cause_mortality_coconino_military_count

# read death by suicide data with categorical data lumps
azdhs_suicide_data_extract_plus_categorical_lumps <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract_plus_categorical_lumps")

# read municipal boundaries 
gis_coconino_municipal_boundaries <- suicide_data %>%
  pin_read("gis_coconino_municipal_boundaries")

# color palette: Coconino 
coco_color_palette <- suicide_data %>%
  pin_read("color_palette_coco_4")

# color palette: Gray 
color_palette_gray_7 <- suicide_data %>%
  pin_read("color_palette_gray_7")
```

## Locations of Injury Address (`r years_of_analysis`) 

```{r 02-locations}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- gis_rate_zipcode_spatial %>%
  ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  # geom_sf(mapping = aes(fill = rate_per_100k), color = NA) +
  # scale_fill_viridis_c() +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude ),
             shape = 4,
             size = 3,
             alpha = 4/7,
             color = "#545234") +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    fill = "Rate per 100K",
    color = "Rate per 100K",
    caption = str_wrap("", width = 60)
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```

## Age groups by year (`r years_of_analysis`) 

```{r 4-age-quantile-by-year}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(age_group_5_quant_bins) %>%
  group_by(death_book_year, age_group_5_quant_bins) %>%
  count() %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = death_book_year,
      y = n,
      fill = age_group_5_quant_bins
        ),
    position = "dodge") +
  scale_fill_viridis_d() +
  labs(
    title = "Death by Suicide Grouped According to Age Group Quantiles",
    subtitle = years_of_analysis,
    x = "Year",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```

## Death City

```{r 5-injury-by-city}
# preparation
city_count_ul <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(death_city_name, min = 6)) %>%
  count(city_code) %>%
  summarise(count_ul = (max(n)) * 1.3) %>%
  as.numeric()

city_count_caption <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(death_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code == "Other")

city_count_caption <- str_c("Not shown are",
  as.character(city_count_caption[1, 2]),
  "deaths by suicide from other cities because the count for that city is less than 6.",
  sep = " "
)

city_count_title <- str_c("City of Location Injury Associated with Death by Suicide (",
  years_of_analysis,
  ")",
  sep = ""
)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(death_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code != "Other") %>%
  ggplot() +
  geom_col(mapping = aes(x = reorder(city_code, n), y = n), fill = "#979797") +
  ylim(0, city_count_ul) +
  coord_flip() +
  labs(
    title = str_wrap(city_count_title, width = 100),
    subtitle = str_wrap("Coconino County cities shown only if 6 or more deaths by suicide ", width = 100),
    x = "City",
    y = "Number of deaths",
    caption = str_wrap(city_count_caption,
      width = 100
    )
  ) +
  theme_suicide_data_report
```

## Military Service 

```{r 6-suicide-by-military}
azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident",
    military_service == "yes"
  ) %>% # county residents only %>%
  group_by(death_book_year, military_service) %>%
  count() %>%
  ungroup() %>%
  full_join(azdhs_all_cause_mortality_coconino_military_count) %>%
  mutate(military_service = case_when(
    is.na(military_service) ~ "total",
    TRUE ~ as.character(military_service)
  )) %>%
  pivot_wider(
    names_from = military_service,
    values_from = n
  ) %>%
  drop_na(yes) %>%
  mutate(percent = yes/total) %>%
  ggplot() +
  geom_col(
    mapping = aes(x = death_book_year,
                  y = percent),
    fill = "#979797"
  ) +
  scale_y_continuous(
    labels = percent,
    # limits  = c(0, .05)
    ) +
  labs(
    title = "Death by Suicide Related to Military Service as a Proportion of All Deaths",
    subtitle = years_of_analysis,
    x = "Year",
    y = "Percentage",
    caption = "Numerator is death by suicide related to military service. Denominator is count of all Coconino deaths."
  ) +
  theme_suicide_data_report
```

## Locations of Injury Address by Method (`r years_of_analysis`) 

```{r 07-location-by-method}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide Grouped by Method (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude,
                           group = method_code_cdc_lump,
                           color = method_code_cdc_lump,
                           shape = method_code_cdc_lump),
             size = 2,
             alpha = 4/7) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    caption = str_wrap("", width = 60),
    shape = "Method",
    color = "Method"
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```

## Locations of Injury Address by Race (`r years_of_analysis`) 

```{r 08-location-by-race}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide Grouped by Race (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude,
                           group = race_code_lump,
                           color = race_code_lump,
                           shape = race_code_lump),
             size = 2,
             alpha = 4/7) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    caption = str_wrap("", width = 60),
    color = "Race",
    shape = "Race"
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```

## Locations of Injury Address by Age Group (`r years_of_analysis`) 

```{r 09-location-by-age}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide Grouped by Age (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude,
                           group = age_group_5_quant_bins,
                           color = age_group_5_quant_bins,
                           shape = age_group_5_quant_bins),
             size = 2,
             alpha = 4/7) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    caption = str_wrap("", width = 60),
    color = "Age group",
    shape = "Age group"
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```

## Locations of Injury Address by County Resident (`r years_of_analysis`) 

```{r 10-location-by-residence}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide Grouped by County Residence (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude,
                           group = county_resident,
                           color = county_resident,
                           shape = county_resident),
             size = 2,
             alpha = 4/7) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    caption = str_wrap("", width = 60),
    shape = "County Residence",
    color = "County Residence"
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```

## Locations of Injury Address by Sex (`r years_of_analysis`) 

```{r 11-location-by-sex}
geo_plot_title <- str_c("Locations of Injury Associated with Death by Suicide Grouped by Sex (",
  years_of_analysis,
  ")",
  sep = ""
)

geo_address_list_points <- geocoded_injury_location_spatial %>%
  filter(longitude >= -113.3542 & longitude < -110.7507) %>%
  filter(latitude > 34.25841 & latitude < 37.0031) %>%
  filter(str_detect(injaddr_county, "coconino"))

death_by_suicide_rate_geospatial <- ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  geom_sf_text(data = gis_coconino_municipal_boundaries,
                mapping = aes(label = NAME),
                alpha = 4/7,
               fontface = "bold") +
  geom_jitter(data = geo_address_list_points,
             mapping = aes(x = longitude,
                           y = latitude,
                           group = sex,
                           color = sex,
                           shape = sex),
             size = 2,
             alpha = 4/7) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 100),
    subtitle = str_wrap("Including only those locations where geocoded data exists.",
      width = 100
    ),
    caption = str_wrap("", width = 60),
    color = "Sex",
    shape = "Sex"
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.4,
    height = 0.4
  )
```


## Percent of death by suicide compared to total number of deaths  (`r years_of_analysis`) 


```{r 12-percent-of-total}
count_of_suicide_by_year <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(county_resident == "Resident") %>%
  count(death_book_year)

suicide_as_percent_5_year_mean <- full_join(
  x = azdhs_all_cause_mortality_coconino_resident_count,
  y = count_of_suicide_by_year,
  by = "death_book_year"
) %>%
  filter(death_book_year %in% analysis_year_range) %>%
  summarise(n.x = sum(n.x),
            n.y = sum(n.y)) %>%
  transmute(percent = n.y/n.x) %>%
  as.numeric()

full_join(
  x = azdhs_all_cause_mortality_coconino_resident_count,
  y = count_of_suicide_by_year,
  by = "death_book_year"
) %>%
  filter(death_book_year %in% analysis_year_range) %>%
  mutate(percent = n.y/n.x) %>%
  ggplot() +
  geom_col(mapping = aes(
    x = death_book_year,
    y = percent
  ),
  fill = "#979797") +
  scale_y_continuous(
    labels = percent,
    limits  = c(0, .1)
  ) +
  geom_hline(mapping = aes(
    yintercept = suicide_as_percent_5_year_mean,
    color = "5 year average"),
    size = 1.5,
    alpha = 2/3
  ) +
  scale_color_manual(values = "#8CBDB8") +
  labs(
    title = "Suicide as a Percentage of Total Deaths",
    subtitle = "Coconino Residents",
    x = "Year",
    y = "Percent",
    color = NULL
  ) +
  theme_suicide_data_report
```


## By Marital Status (`r years_of_analysis`) 

```{r 13-marital}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(marital_code) %>%
  group_by(marital_code) %>%
  count(sort = TRUE) %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = fct_reorder(marital_code, n),
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  coord_flip() +
  labs(
    title = "Death by Suicide Grouped According to Marital Status",
    subtitle = years_of_analysis,
    x = "Marital Status",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```

## By Educational Attainment (`r years_of_analysis`) 

```{r 14-edu}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(edu_code) %>%
  group_by(edu_code) %>%
  count(sort = TRUE) %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = fct_reorder(edu_code, n),
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  coord_flip() +
  labs(
    title = "Death by Suicide Grouped According to Educational Attainment",
    subtitle = years_of_analysis,
    x = "Educational Attainment",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```


## By Day of the Week (`r years_of_analysis`) 

```{r 15-weekday}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(date_of_death_day_week) %>%
  group_by(date_of_death_day_week) %>%
  count() %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = date_of_death_day_week,
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  labs(
    title = "Death by Suicide Grouped According to Day of the Week",
    subtitle = years_of_analysis,
    x = "Day of the Week",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```

## By Day of the Month (`r years_of_analysis`) 

```{r 16-day-of-month}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(date_of_death_day_month) %>%
  group_by(date_of_death_day_month) %>%
  count() %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = date_of_death_day_month,
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  scale_x_continuous(breaks = seq(1,31, 3)) +
  labs(
    title = "Death by Suicide Grouped According to Day of the Month",
    subtitle = years_of_analysis,
    x = "Day",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```


## By Month (`r years_of_analysis`) 

```{r 17-month}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(date_of_death_month) %>%
  group_by(date_of_death_month) %>%
  count() %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = date_of_death_month,
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  scale_x_continuous(breaks = seq(1,12, 3), labels = c("Jan", "Apr", "Jul", "Oct")) +
  labs(
    title = "Death by Suicide Grouped According to Month",
    subtitle = years_of_analysis,
    x = "Month",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = ""
  ) +
  theme_suicide_data_report
```


## By Birth Country (`r years_of_analysis`) 

```{r 18-birth-country}
# glimpse(azdhs_suicide_data_extract_plus_categorical_lumps)

azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(death_book_year %in% analysis_year_range) %>%
  drop_na(birth_country) %>%
  group_by(birth_country) %>%
  count() %>%
  filter(birth_country != "United States" & birth_country != "Unknown") %>%
  ggplot() +
  geom_col(
    mapping = aes(
      x = reorder(birth_country, n),
      y = n
        ),
    position = "dodge",
    fill = "#979797") +
  coord_flip() +
  labs(
    title = "Death by Suicide Grouped According to Country of Birth",
    subtitle = years_of_analysis,
    x = "Birth Country",
    y = "Number of Deaths",
    fill = "Age Group",
    caption = "Excluding where the country of birth is the United States"
  ) +
  theme_suicide_data_report
```
