---
title: "Coconino County Health and Human Services Vital Record and Medical Examiner Data for Death by Suicide"
subtitle: ""
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  beamer_presentation:
    colortheme: "seahorse"
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  fig.width = 11
)
```

```{r libraries}
library(here)
library(tidyverse)
library(pins)
library(knitr)
library(RColorBrewer)
library(scales)
library(cowplot)
```

```{r theme}
theme_suicide_data_report <- theme_classic() +
  theme(
    axis.line = element_line(size = .5, color = "#979797"),
    axis.ticks = element_line(size = .5, color = "#979797"),
    aspect.ratio = 3 / 5.5,
    text = element_text(
      size = 14,
      color = "#3b3b3b"
    )
  )
```

```{r read}
# load pin board
suicide_data <- board_folder("S:/HIPAA Compliance/SAS Files/Coconino Deaths/Suicide/data-raw")

# read complete data set
azdhs_suicide_data_extract <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract")

# years of analysis
analysis_year_range <- c(as.character(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_current)))

# lower limit
analysis_year_range_ll <- as.character(min(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# upper limit
analysis_year_range_ul <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# current year
analysis_year_current <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_current)))

# create a character string showing range of years included in report
years_of_analysis <- str_c(analysis_year_range_ll,
  analysis_year_range_ul,
  sep = " - "
)

# read count of death by suicide including ME pending cases
azdhs_death_by_suicide_coco_resident_me <- suicide_data %>%
  pin_read("azdhs_death_by_suicide_coco_resident_me")

# read death by suicide data
death_by_suicide_coconino_resident_age_adjusted_rate <- suicide_data %>%
  pin_read("death_by_suicide_coconino_resident_age_adjusted_rate")

# read comparison of age adjusted rates
rates_usa_az_coco <- suicide_data %>%
  pin_read("death_by_suicide_rate_comparison")

# read geospatial rate by zip code
gis_rate_zipcode_spatial <- suicide_data %>%
  pin_read("gis_rate_zipcode_spatial")

# read geospatial coconino county boundary
gis_coco_county_spatial <- suicide_data %>%
  pin_read("gis_coco_county_spatial")

# read geospatial coconino county primary roads
gis_coco_county_roads_spatial <- suicide_data %>%
  pin_read("gis_coco_county_roads_spatial")

# read data by race
death_by_suicide_coconino_resident_race_specific_rate <- suicide_data %>%
  pin_read("death_by_suicide_coconino_resident_race_specific_rate")

# read population denominators by race
azdhs_pop_data_by_sex_gender <- suicide_data %>%
  pin_read("azdhs_pop_data_by_sex_gender")

# read death by suicide data with categorical data lumps
azdhs_suicide_data_extract_plus_categorical_lumps <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract_plus_categorical_lumps")

# most recent death
most_recent <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  summarise(max(date_of_death))

most_recent <- format(most_recent$`max(date_of_death)`, "%d %B %Y")
```

## Confidentiality


**Please do not distribute.**


**This information is confidential and intended for the purposes of this group.**


## Notes


- All data presented for year `r analysis_year_range_ul` and `r analysis_year_current` is subject to change.
- Mortality data is provided by AZDHS.
- Pending cases with the medical examiner's office are included only in the number of deaths (slide 4). 
- AZDHS data is up to date as of `r most_recent`.


## Number of Deaths by Suicide Among Coconino County Residents (`r analysis_year_range_ll` - `r analysis_year_current`)


```{r 04-number}
# preparation
# plot upper limits
death_by_suicide_count_ul <- azdhs_death_by_suicide_coco_resident_me %>%
  summarise(count_ul = (max(n)) * 1.3) %>%
  as.numeric()

# title, subitle, & caption
count_me_pending_cases <- azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    me_brief == "yes"
  ) %>%
  count() %>%
  as.numeric()

plot_caption <- str_c(
  "2021 & 2022 figures are based off preliminary death data and include ",
  count_me_pending_cases,
  " possible deaths by suicide currently under investigation by the Coconino County Medical Examiner's Office."
)

# plot
azdhs_death_by_suicide_coco_resident_me %>%
  ggplot(mapping = aes(x = death_book_year, y = n, group = 1)) +
  geom_line(color = "#0476D0", size = 1) +
  geom_label(aes(label = n),
    alpha = (5 / 6),
    color = "white",
    fill = "#0476D0",
    fontface = "bold"
  ) +
  ylim(0, death_by_suicide_count_ul) +
  labs(
    title = str_wrap("Total Number of Deaths by Suicide for Each Year",
      width = 120
    ),
    subtitle = str_wrap("Includes only Coconino County residents.", width = 120),
    x = "Year",
    y = "Number of deaths",
    caption = str_wrap(plot_caption,
      width = 120
    )
  ) +
  theme_suicide_data_report
```


## Comparison of Age-Adjusted Death by Suicide Rate with Coconino County, Arizona, and the U.S. (`r years_of_analysis`)


```{r 05-rate-comparison}
# preparation
age_adj_rates_ul <- rates_usa_az_coco %>%
  summarise(rate_ul = (max(age_adjusted_rate)) * 1.3) %>%
  as.numeric()

# plot
rates_usa_az_coco %>%
  filter(age_adjusted_rate > 0) %>%
  ggplot() +
  geom_line(
    mapping = aes(
      x = year,
      y = age_adjusted_rate,
      group = geography,
      color = geography
    ),
    size = 1,
    alpha = (5 / 6)
  ) +
  geom_label(
    mapping = aes(
      x = year,
      y = age_adjusted_rate,
      label = round(age_adjusted_rate, 1),
      fill = geography
    ),
    color = "white",
    fontface = "bold",
    alpha = (5 / 6)
  ) +
  ylim(0, age_adj_rates_ul) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = str_wrap("Age-Adjusted Rate of Death by Suicide", width = 120),
    subtitle = str_wrap("Includes only Coconino County residents.", width = 120),
    x = "Year",
    y = "Age adjusted rate per 100K",
    color = "",
    fill = "",
    caption = str_wrap("2020 & 2021 death by suicide rate is not yet available for Arizona and the United States. The Coconino rate is derived from the population denominator provided by AZDHS with year 2020 population estimate being carried over to 2021.", width = 120)
  ) +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  ) +
  theme_suicide_data_report
```

## Death by Suicide in Coconino County by Coconino County Residence (`r years_of_analysis`)

```{r 06-resident}
# preparation
count_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(death_book_year %in% analysis_year_range) %>%
  group_by(death_book_year, county_resident) %>% # county residents only
  count() %>%
  ungroup() %>%
  summarise(count_ul = (max(n)) * 1.3) %>%
  as.numeric()

# plot
azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(death_book_year %in% analysis_year_range) %>%
  group_by(death_book_year, county_resident) %>% # county residents only
  count() %>%
  ungroup() %>%
  ggplot(mapping = aes(
    x = death_book_year,
    y = n,
    group = county_resident,
    color = county_resident
  )) +
  geom_line(size = 1, alpha = (5 / 6)) +
  geom_label(
    mapping = aes(
      fill = county_resident,
      x = death_book_year,
      y = n,
      label = (n)
    ),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  ylim(0, count_ul) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Death by Suicide Occurring in Coconino County",
    x = "Year",
    y = "Number of deaths",
    color = "Coconino Resident",
    fill = "Coconino Resident",
    caption = "Death by suicide for non-residents occurred within Coconino County."
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Residence of Coconino County Suicides (`r years_of_analysis`)


```{r 07-city-list}
# preparation
city_count_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(residence_city_name, min = 6)) %>%
  count(city_code) %>%
  summarise(count_ul = (max(n)) * 1.3) %>%
  as.numeric()

city_count_caption <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(residence_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code == "Other")

city_count_caption <- str_c("Not shown are",
  as.character(city_count_caption[1, 2]),
  "deaths by suicide from other cities because the count for that city is less than 6.",
  sep = " "
)

city_count_title <- str_c("City Residence of Death by Suicide in Coconino County (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(residence_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code != "Other") %>%
  ggplot() +
  geom_col(mapping = aes(x = reorder(city_code, n), y = n), fill = "#647c90") +
  ylim(0, city_count_ul) +
  coord_flip() +
  geom_text(mapping = aes(x = reorder(city_code, n), y = n, label = (n)), nudge_y = 4, color = "#647c90", fontface = "bold") +
  labs(
    title = str_wrap(city_count_title, width = 120),
    subtitle = str_wrap("Cities shown only if 6 or more deaths by suicide. Includes only Coconino County residents.", width = 120),
    x = "City",
    y = "Number of deaths",
    caption = str_wrap(city_count_caption,
      width = 120
    )
  ) +
  theme_suicide_data_report
```

## Rate of Death by Suicide by Zip Code for Coconino County Residents (`r years_of_analysis`)

```{r 08-geospatial}
geo_plot_title <- str_c("Death by Suicide for Coconino County Residents (",
  years_of_analysis,
  ")",
  sep = ""
)

death_by_suicide_rate_geospatial <- gis_rate_zipcode_spatial %>%
  ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(mapping = aes(fill = rate_per_100k), color = NA) +
  scale_fill_viridis_c() +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  # geom_sf_label(data = coconino_municipal,
  #              mapping = aes(label = NAME),
  #              nudge_y = .05,
  #              nudge_x = .05,
  #              size = 2,
  #              alpha = (1/3)
  #              ) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 120),
    subtitle = str_wrap("Including only those zip codes with more than 6 deaths by suicide of Coconino County Residents.",
      width = 120
    ),
    fill = "Rate per 100K",
    color = "Rate per 100K",
    caption = str_wrap("Coconino rate derived with population denominator provided by AZDHS. Year 2020 population estimate carried over to 2021.", width = 60)
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.425,
    height = 0.425
  )
```

## Age Groups of Death by Suicide for Coconino County Residents (`r years_of_analysis`)

```{r 09-age-group}
# preparation
death_by_age_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(age_group_6cat) %>%
  count() %>%
  ungroup() %>%
  summarise(age_ul = (max(n) * 1.3)) %>%
  as.numeric()

death_by_age_title <- str_c("Death by Suicide for Each Age Group (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(age_group_6cat) %>%
  count() %>%
  ungroup() %>%
  ggplot() +
  geom_bar(
    mapping = aes(x = age_group_6cat, y = n),
    stat = "identity", fill = "#01949A"
  ) +
  ylim(0, death_by_age_ul) +
  coord_flip() +
  geom_text(
    mapping = aes(
      x = age_group_6cat,
      y = n,
      label = (n)
    ),
    nudge_y = 4,
    color = "#808080",
    fontface = "bold"
  ) +
  # annotate(
  #   geom = "text", x = 1, y = 80,
  #   label = str_wrap("2021 figures are based off preliminary death data", width = 40), hjust = 0, vjust = 1
  # ) +
  labs(
    title = str_wrap(death_by_age_title,
      width = 120
    ),
    subtitle = "Including only Coconino County residents.",
    x = "Age Group",
    y = "Number of deaths"
  ) +
  theme_suicide_data_report
```

## Age Groups for Death by Suicide (`r analysis_year_range_ul`)

```{r 10-age-group-current}
# preparation
age_year_title <- str_c("Death by Suicide for Each Age Group, Year to Date (",
  analysis_year_range_ul,
  ")",
  sep = ""
)

count_me_pending_cases <- azdhs_suicide_data_extract %>%
  filter(
    death_book_year == analysis_year_range_ul,
    me_brief == "yes"
  ) %>%
  count() %>%
  as.numeric()

# plot_caption <- str_c(analysis_year_range_ul,
#                       " figures are based off preliminary death data and include ",
#                       count_me_pending_cases,
#                       " possible deaths by suicide currently under investigation by the Coconino County Medical Examiner's Office.")

count_age_ul <- azdhs_suicide_data_extract %>%
  filter(
    death_book_year == analysis_year_range_ul,
    county_resident == "Resident"
  ) %>% # county residents only
  count(age_group_6cat) %>%
  summarise(count_age_ul = (max(n) * 1.5)) %>%
  as.numeric()

# plot
azdhs_suicide_data_extract %>%
  filter(
    death_book_year == analysis_year_range_ul,
    county_resident == "Resident"
  ) %>% # county residents only
  count(age_group_6cat) %>%
  ggplot(mapping = aes(x = age_group_6cat, y = n)) +
  geom_bar(stat = "identity", fill = "#01949A") +
  geom_text(
    mapping = aes(
      x = age_group_6cat,
      y = n,
      label = (n)
    ),
    nudge_y = .5,
    color = "#808080",
    fontface = "bold"
  ) +
  ylim(0, count_age_ul) +
  coord_flip() +
  labs(
    title = str_wrap(age_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Age Group",
    y = "Number of deaths",
    caption = ""
  ) +
  theme_suicide_data_report
```

## Age Groups of Death by Suicides for Coconino County Residents (`r years_of_analysis`)


```{r 11-age-group-by-year}
age_group_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year, age_group_6cat) %>%
  count(age_group_6cat) %>%
  ungroup() %>%
  summarise(age_group_ul = (max(n) * 1.3)) %>%
  as.numeric()

age_group_by_year_title <- str_c("Age Groups of Death by Suicide (",
  years_of_analysis,
  ")",
  sep = ""
)

azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year, age_group_6cat) %>%
  count(age_group_6cat) %>%
  # mutate(year = as.integer(death_book_year)) %>%
  ggplot(mapping = aes(x = death_book_year, y = n)) +
  geom_bar(
    position = "dodge",
    stat = "identity",
    mapping = aes(fill = age_group_6cat),
    # color = "#979797"
  ) +
  scale_fill_brewer(palette = "Set3", name = "Age Group") +
  geom_text(
    position = position_dodge(width = .9),
    mapping = aes(label = (n), group = age_group_6cat),
    color = "#3b3b3b",
    vjust = -.5
  ) +
  labs(
    title = str_wrap(age_group_by_year_title,
      width = 120
    ),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths"
  ) +
  ylim(0, age_group_ul) +
  scale_x_discrete(breaks = seq(as.numeric(analysis_year_range_ll), as.numeric(analysis_year_range_ul), 1)) +
  theme_suicide_data_report
```

## Death by Suicide and Race for Coconino County Residents (`r years_of_analysis`)

```{r 12-race-by-year}
# preparation
# color palette
# set color palettes
coconino_population_by_race_palette <- death_by_suicide_coconino_resident_race_specific_rate$race_code_lump

death_by_suicide <- azdhs_suicide_data_extract_plus_categorical_lumps

palette_race_4 <- brewer.pal(n = 4, name = "Set2")
names(palette_race_4) <- levels(factor(c(
  levels(death_by_suicide$race_code_lump),
  levels(coconino_population_by_race_palette)
))) # Extract all levels of both data

race_by_year_title <- str_c("Death by Suicide and Race (",
  years_of_analysis,
  ")",
  sep = ""
)

race_count_ul <- death_by_suicide %>%
  group_by(death_book_year) %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  count(race_code_lump) %>%
  ungroup() %>%
  summarise(count_ul = (max(n) * 1.3)) %>%
  as.numeric()

death_by_race_year_count <- death_by_suicide %>%
  group_by(death_book_year) %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  count(race_code_lump) %>%
  pivot_wider(
    names_from = race_code_lump, values_from = n
  ) %>%
  ungroup()

death_by_race_year_count[is.na(death_by_race_year_count)] <- 0

# plot
death_by_race_year_count %>%
  pivot_longer(
    cols = 2:5,
    names_to = "race_code_lump",
    values_to = "n"
  ) %>%
  ggplot(mapping = aes(
    x = death_book_year,
    y = n,
    group = race_code_lump,
    color = race_code_lump
  )) +
  geom_line(size = 1) +
  geom_label(
    mapping = aes(
      fill = race_code_lump,
      x = death_book_year,
      y = n,
      label = (n)
    ),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  scale_color_manual(values = palette_race_4) +
  scale_fill_manual(values = palette_race_4) +
  ylim(0, race_count_ul) +
  labs(
    title = str_wrap(race_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    color = "Race",
    fill = "Race"
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```


## Percentage of Death by Suicide by Race (`r years_of_analysis`)


```{r 13-race-by-percent}
# preparation
# Percent stacked barchart
death_by_suicide_race_percentage <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  count(race_code_lump) %>%
  mutate(n_percent = round(100 * (n / sum(n)), digits = 0)) %>%
  mutate(group = "Coconino Death by Suicide") %>%
  mutate(race_code = race_code_lump)

coconino_population_by_race_percentage <- azdhs_pop_data_by_sex_gender %>%
  filter(race_ethnicity != "All groups") %>%
  filter(year == 2020) %>%
  filter(age_group == "total") %>%
  filter(sex == "Total") %>%
  filter(area == "Coconino") %>%
  mutate(
    estimate = as.numeric(estimate),
    race_code_lump =
      case_when(
        race_ethnicity == "White non-Hispanic" ~ "White Non-Hispanic",
        race_ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino (any race)",
        race_ethnicity == "Black or African American" ~ "Other",
        race_ethnicity == "Asian or Pacific Islander" ~ "Other",
        race_ethnicity == "American Indian or Alaska Native" ~ "American Indian and Alaska Native"
      )
  ) %>%
  group_by(race_code_lump) %>%
  summarise(estimate = sum(estimate)) %>%
  mutate(
    n_percent = round(100 * (estimate / sum(estimate)), digits = 0),
    group = "Coconino 2020 Population"
  )

suicide_by_race_percentage <- bind_rows(
  death_by_suicide_race_percentage,
  coconino_population_by_race_percentage
)

suicide_by_race_percentage <- full_join(
  x = death_by_suicide_race_percentage,
  y = coconino_population_by_race_percentage,
  by = c(
    "n" = "estimate",
    "race_code_lump" = "race_code_lump",
    "n_percent" = "n_percent",
    "group" = "group"
  )
)

percent_race_by_year_title <- str_c("Percentage of Death by Suicide by Race in Comparison to 2020 Population (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
suicide_by_race_percentage %>%
  ggplot(mapping = aes(fill = race_code_lump, y = n_percent, x = group)) +
  geom_col(position = "fill") +
  geom_text(
    position = position_fill(vjust = 0.5),
    mapping = aes(label = comma(n)),
    color = "white"
  ) +
  scale_fill_manual(values = palette_race_4) +
  # scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = percent) +
  labs(
    title = str_wrap(percent_race_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "",
    y = "Percentage",
    fill = "Race",
    caption = str_wrap("The values presented within the plot represent the Coconino 2020 Population estimate and the number of deaths by suicide, respectively.", width = 120)
  ) +
  theme_suicide_data_report
```

## Deaths by Suicide and Gender for Coconino County Residents (`r years_of_analysis`)

```{r 14-sex-by-year}
# preparation
sex_count_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(sex) %>%
  ungroup() %>%
  summarise(sex_count_ul = (max(n) * 1.3)) %>%
  as.numeric()

gender_by_year_title <- str_c("Deaths by Suicide by Sex (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(sex) %>%
  ggplot(mapping = aes(x = death_book_year, y = n, group = sex, color = sex)) +
  geom_line(size = 1) +
  geom_label(
    mapping = aes(fill = sex, x = death_book_year, y = n, label = (n)),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  ylim(0, sex_count_ul) +
  scale_color_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  labs(
    title = str_wrap(gender_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    color = "Sex",
    fill = "Sex"
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Total Deaths by Suicide and Gender for Coconino County Residents (`r years_of_analysis`)

```{r 15-sex}
# preparation
by_sex_total_count_ul <- azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count() %>%
  ungroup() %>%
  summarise(count_ul = (max(n)) * 1.3) %>%
  as.numeric()

gender_title <- str_c("Death by Suicide for Each Sex (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count() %>%
  ungroup() %>%
  ggplot(mapping = aes(x = sex, y = n)) +
  geom_col(mapping = aes(fill = sex), show.legend = FALSE) +
  geom_text(
    mapping = aes(label = n, y = n),
    color = "#525252",
    nudge_y = 8
  ) +
  scale_fill_manual(values = c("#7FC97F", "#BEAED4")) +
  scale_y_continuous(limits = c(0, by_sex_total_count_ul)) +
  labs(
    title = str_wrap(gender_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    fill = "Sex"
  ) +
  theme_suicide_data_report
```

## Methods of Suicides in Coconino County (`r years_of_analysis`)

```{r 16-method}
# preparation
method_count_ul <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  summarise(count_ul = (max(n) * 1.3)) %>%
  as.numeric()

method_title <- str_c("Methods of Suicides (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  ggplot(mapping = aes(x = death_book_year, y = n, fill = method_code_cdc_lump)) +
  geom_bar(position = "dodge", stat = "identity") +
  ylim(0, method_count_ul) +
  geom_text(
    position = position_dodge(width = 1),
    mapping = aes(label = (n)),
    color = "#3b3b3b",
    vjust = -1
  ) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
    title = str_wrap(method_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    fill = "Method",
    caption = str_wrap(
      "Method is coded from a free text variable and is an injury description.",
      width = 120
    )
  ) +
  theme_suicide_data_report
```

## Methods of Suicide by Gender in Coconino County (`r years_of_analysis`)

```{r 17-method-by-sex}
# preparation
method_count_ul <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  summarise(count_ul = (max(n) * 1.3)) %>%
  as.numeric()

method_by_sex_title <- str_c("Methods of Suicides by Sex (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(me_brief == "no") %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  ggplot(mapping = aes(x = sex, y = n, group = method_code_cdc_lump, fill = method_code_cdc_lump)) +
  geom_bar(position = "dodge", stat = "identity") +
  ylim(0, method_count_ul) +
  geom_text(
    position = position_dodge(width = 1),
    mapping = aes(label = (n)),
    color = "#3b3b3b",
    vjust = -1
  ) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
    title = str_wrap(method_by_sex_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Sex",
    y = "Number of deaths",
    fill = "Method",
    caption = str_wrap(
      "Method is coded from a free text variable and is an injury description.",
      width = 120
    )
  ) +
  theme_suicide_data_report
```
