---
title: ""
subtitle: ""
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  beamer_presentation:
    colortheme: "seahorse"
    slide_level: 2

params:
  username:
    label: "NSSP Username: "
    value: ""
    input: text
    placeholder: "username"
  password:
    label: "NSSP Password: "
    value: ""
    input: password
    placeholder: "password"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  fig.width = 11
)
```

```{r libraries}
library(here)
library(tidyverse)
library(pins)
library(knitr)
library(RColorBrewer)
library(scales)
library(cowplot)
library(Rnssp)
```

```{r theme}
theme_suicide_data_report <- theme_classic() +
  theme(
    axis.line = element_line(size = .5, color = "#979797"),
    axis.ticks = element_line(size = .5, color = "#979797"),
    aspect.ratio = 3 / 5.5,
    text = element_text(
      size = 14,
      color = "#3b3b3b"
    )
  )
```

```{r read}
# load pin board
suicide_data <- board_folder("")

# read complete data set
azdhs_suicide_data_extract <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract")

# years of analysis
analysis_year_range <- c(as.character(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# lower limit
analysis_year_range_ll <- as.character(min(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# upper limit
analysis_year_range_ul <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_ll):unique(azdhs_suicide_data_extract$year_analysis_ul)))

# current year
analysis_year_current <- as.character(max(unique(azdhs_suicide_data_extract$year_analysis_current)))

# create a character string showing range of years included in report
years_of_analysis <- str_c(analysis_year_range_ll,
  analysis_year_range_ul,
  sep = " - "
)

# read count of death by suicide including ME pending cases
azdhs_death_by_suicide_coco_resident_me <- suicide_data %>%
  pin_read("number_of_deaths_by_suicide_by_year")

# read death by suicide data
death_by_suicide_coconino_resident_age_adjusted_rate <- suicide_data %>%
  pin_read("death_by_suicide_coconino_resident_age_adjusted_rate")

# read comparison of age adjusted rates
rates_usa_az_coco <- suicide_data %>%
  pin_read("death_by_suicide_rate_comparison")

# read geospatial rate by zip code
gis_rate_zipcode_spatial <- suicide_data %>%
  pin_read("gis_rate_zipcode_spatial")

# read geospatial coconino county boundary
gis_coco_county_spatial <- suicide_data %>%
  pin_read("gis_coco_county_spatial")

# read geospatial coconino county primary roads
gis_coco_county_roads_spatial <- suicide_data %>%
  pin_read("gis_coco_county_roads_spatial")

# read data by race
death_by_suicide_coconino_resident_race_specific_rate <- suicide_data %>%
  pin_read("death_by_suicide_coconino_resident_race_specific_rate")

# read population denominators by race
azdhs_pop_data_by_sex_gender <- suicide_data %>%
  pin_read("azdhs_pop_data_by_sex_gender")

# read death by suicide data with categorical data lumps
azdhs_suicide_data_extract_plus_categorical_lumps <- suicide_data %>%
  pin_read("azdhs_suicide_data_extract_plus_categorical_lumps")

# medical examiner report
me_brief_count <- suicide_data %>%
  pin_read("me_brief_count")

# suicide in 2022 plus medical examiner report
suicide_data_coco_2022 <- suicide_data %>%
  pin_read("suicide_data_coco_2022")

# most recent death
most_recent <- azdhs_suicide_data_extract %>%
  summarise(max(date_of_death))

most_recent <- format(most_recent$`max(date_of_death)`, "%d %B %Y")
```

## Confidentiality


**Please do not distribute.**


**This information is confidential and intended for the purposes of this group.**


## Notes


- All data presented for year `r analysis_year_range_ul` and `r analysis_year_current` is subject to change.
- Mortality data is provided by AZDHS.
- Pending cases with the medical examiner's office are included only in the number of deaths (slide 4) and the count of deaths by age group in the year, `r analysis_year_current` (slide 10).
- AZDHS data is up to date as of `r most_recent`.
- Syndromic surveillance (slide 19) provides public health officials with a timely system for detecting, understanding, and monitoring health events.


## Number of Deaths by Suicide Among Coconino County Residents (`r analysis_year_range_ll` - `r analysis_year_current`)


```{r 04-number}
# preparation
me_brief_count_caption <- me_brief_count %>%
  filter(death_book_year == "2022")

# title, subtitle, & caption
plot_caption <- str_c(
  "2021 & 2022 figures are based off preliminary death data and include ",
  me_brief_count_caption$n,
  " possible deaths by suicide currently under investigation by the Coconino County Medical Examiner's Office."
)

# plot
azdhs_death_by_suicide_coco_resident_me %>%
  ggplot(mapping = aes(x = death_book_year, y = n, group = 1)) +
  geom_line(color = "#0476D0", size = 1) +
  geom_label(aes(label = n),
    alpha = (5 / 6),
    color = "white",
    fill = "#0476D0",
    fontface = "bold"
  ) +
  ylim(0, NA) +
  labs(
    title = str_wrap("Total Number of Deaths by Suicide for Each Year",
      width = 120
    ),
    subtitle = str_wrap("Includes only Coconino County residents.", width = 120),
    x = "Year",
    y = "Number of deaths",
    caption = str_wrap(plot_caption,
      width = 120
    )
  ) +
  theme_suicide_data_report
```


## Comparison of Age-Adjusted Death by Suicide Rate with Coconino County, Arizona, and the U.S. (`r years_of_analysis`)


```{r 05-rate-comparison}
# plot
rates_usa_az_coco %>%
  filter(age_adjusted_rate > 0) %>%
  filter(year %in% analysis_year_range) %>%
  ggplot() +
  geom_line(
    mapping = aes(
      x = year,
      y = age_adjusted_rate,
      group = geography,
      color = geography
    ),
    size = 1,
    alpha = (5 / 6)
  ) +
  geom_label(
    mapping = aes(
      x = year,
      y = age_adjusted_rate,
      label = round(age_adjusted_rate, 1),
      fill = geography
    ),
    color = "white",
    fontface = "bold",
    alpha = (5 / 6)
  ) +
  ylim(0, NA) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = str_wrap("Age-Adjusted Rate of Death by Suicide", width = 120),
    subtitle = str_wrap("Includes only Coconino County residents.", width = 120),
    x = "Year",
    y = "Age adjusted rate per 100K",
    color = "",
    fill = "",
    caption = str_wrap("2021 death by suicide rate is not yet available for Arizona and the United States. The Coconino rate is derived from the population denominator provided by AZDHS with year 2020 population estimate being carried over to 2021.", width = 120)
  ) +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  ) +
  theme_suicide_data_report
```

## Death by Suicide in Coconino County by Coconino County Residence (`r years_of_analysis`)

```{r 06-resident}
# plot
azdhs_suicide_data_extract %>%
  filter(death_book_year %in% analysis_year_range) %>%
  group_by(death_book_year, county_resident) %>% # county residents only
  count() %>%
  ungroup() %>%
  ggplot(mapping = aes(
    x = death_book_year,
    y = n,
    group = county_resident,
    color = county_resident
  )) +
  geom_line(size = 1, alpha = (5 / 6)) +
  geom_label(
    mapping = aes(
      fill = county_resident,
      x = death_book_year,
      y = n,
      label = (n)
    ),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  ylim(0, NA) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Death by Suicide Occurring in Coconino County",
    x = "Year",
    y = "Number of deaths",
    color = "Coconino Resident",
    fill = "Coconino Resident",
    caption = "Death by suicide for non-residents occurred within Coconino County."
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Residence of Coconino County Suicides (`r years_of_analysis`)


```{r 07-city-list}
# preparation
city_count_caption <- azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>%
  mutate(city_code = fct_lump_min(residence_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code == "Other")

city_count_caption <- str_c("Not shown are",
  as.character(city_count_caption$n),
  "deaths by suicide from other cities because the count for that city is less than 6.",
  sep = " "
)

city_count_title <- str_c("City Residence of Death by Suicide in Coconino County (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  mutate(city_code = fct_lump_min(residence_city_name, min = 6)) %>%
  count(city_code) %>%
  filter(city_code != "Other") %>%
  ggplot() +
  geom_col(mapping = aes(x = reorder(city_code, n), y = n), fill = "#647c90") +
  ylim(0, NA) +
  coord_flip() +
  geom_text(mapping = aes(x = reorder(city_code, n), y = n, label = (n)), nudge_y = 4, color = "#647c90", fontface = "bold") +
  labs(
    title = str_wrap(city_count_title, width = 120),
    subtitle = str_wrap("Cities shown only if 6 or more deaths by suicide. Includes only Coconino County residents.", width = 120),
    x = "City",
    y = "Number of deaths",
    caption = str_wrap(city_count_caption,
      width = 120
    )
  ) +
  theme_suicide_data_report
```

## Rate of Death by Suicide by Zip Code for Coconino County Residents (`r years_of_analysis`)

```{r 08-geospatial}
geo_plot_title <- str_c("Death by Suicide for Coconino County Residents (",
  years_of_analysis,
  ")",
  sep = ""
)

death_by_suicide_rate_geospatial <- gis_rate_zipcode_spatial %>%
  ggplot() +
  geom_sf(data = gis_coco_county_spatial, fill = "#C5C5C5", color = "#aeaeae") +
  geom_sf(mapping = aes(fill = rate_per_100k), color = NA) +
  scale_fill_viridis_c() +
  geom_sf(data = gis_coco_county_roads_spatial, color = "#979797") +
  # geom_sf_label(data = coconino_municipal,
  #              mapping = aes(label = NAME),
  #              nudge_y = .05,
  #              nudge_x = .05,
  #              size = 2,
  #              alpha = (1/3)
  #              ) +
  theme_void() +
  coord_sf(expand = TRUE) +
  geom_rect(
    xmin = -111.7191,
    ymin = 35.1224,
    xmax = -111.5068,
    ymax = 35.24328,
    fill = NA,
    colour = "black",
    size = 0.5
  ) +
  labs(
    title = str_wrap(geo_plot_title, width = 120),
    subtitle = str_wrap("Including only those zip codes with more than 6 deaths by suicide of Coconino County Residents.",
      width = 120
    ),
    fill = "Rate per 100K",
    color = "Rate per 100K",
    caption = str_wrap("Coconino rate derived with population denominator provided by AZDHS. Year 2020 population estimate carried over to 2021.", width = 60)
  )

death_by_suicide_rate_geospatial %>%
  ggdraw() +
  draw_plot(
    {
      death_by_suicide_rate_geospatial +
        coord_sf(
          xlim = c(-111.7191, -111.5068),
          ylim = c(35.1224, 35.24328),
          expand = FALSE
        ) +
        theme(legend.position = "none") +
        labs(
          title = "Flagstaff",
          subtitle = "",
          caption = ""
        )
    },
    x = 0,
    y = 0.05,
    width = 0.425,
    height = 0.425
  )
```

## Age Groups of Death by Suicide for Coconino County Residents (`r years_of_analysis`)

```{r 09-age-group}
# preparation
death_by_age_title <- str_c("Death by Suicide for Each Age Group (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(age_group_6cat) %>%
  count() %>%
  ungroup() %>%
  ggplot() +
  geom_bar(
    mapping = aes(x = age_group_6cat, y = n),
    stat = "identity", fill = "#01949A"
  ) +
  ylim(0, NA) +
  coord_flip() +
  geom_text(
    mapping = aes(
      x = age_group_6cat,
      y = n,
      label = (n)
    ),
    nudge_y = 4,
    color = "#808080",
    fontface = "bold"
  ) +
  # annotate(
  #   geom = "text", x = 1, y = 80,
  #   label = str_wrap("2021 figures are based off preliminary death data", width = 40), hjust = 0, vjust = 1
  # ) +
  labs(
    title = str_wrap(death_by_age_title,
      width = 120
    ),
    subtitle = "Including only Coconino County residents.",
    x = "Age Group",
    y = "Number of deaths"
  ) +
  theme_suicide_data_report
```

## Age Groups for Death by Suicide (`r max(azdhs_suicide_data_extract$death_book_year)`)

```{r 10-age-group-current}
# preparation
age_year_title <- str_c("Death by Suicide for Each Age Group, Year to Date (",
  analysis_year_current,
  ")",
  sep = ""
)

count_me_pending_cases <- me_brief_count %>%
  filter(death_book_year == "2022") %>%
  summarise(max(n)) %>%
  as.numeric()

total_suicide_this_year <- azdhs_death_by_suicide_coco_resident_me %>%
  filter(death_book_year == "2022") %>%
  summarise(sum(n)) %>%
  as.numeric()

# plot
suicide_data_coco_2022 %>%
  ggplot(mapping = aes(x = age_group_6cat, y = n)) +
  geom_bar(stat = "identity", fill = "#01949A") +
  geom_text(
    mapping = aes(
      x = age_group_6cat,
      y = n,
      label = (n)
    ),
    nudge_y = .5,
    color = "#808080",
    fontface = "bold"
  ) +
  ylim(0, NA) +
  coord_flip() +
  labs(
    title = str_wrap(age_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Age Group",
    y = "Number of deaths",
    caption = str_wrap(str_c(
      "In the year ",
      analysis_year_current,
      " there have been ",
      total_suicide_this_year,
      " deaths by suicide; including ",
      count_me_pending_cases,
      " cases under investigation by the medical examiner, but not yet included in the mortality data provided by ADHS."
    ), width = 120)
  ) +
  theme_suicide_data_report
```

## Age Groups of Death by Suicides for Coconino County Residents (`r years_of_analysis`)


```{r 11-age-group-by-year}
age_group_by_year_title <- str_c("Age Groups of Death by Suicide (",
  years_of_analysis,
  ")",
  sep = ""
)

azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year, age_group_6cat) %>%
  count(age_group_6cat) %>%
  # mutate(year = as.integer(death_book_year)) %>%
  ggplot(mapping = aes(x = death_book_year, y = n)) +
  geom_bar(
    position = "dodge",
    stat = "identity",
    mapping = aes(fill = age_group_6cat),
    # color = "#979797"
  ) +
  scale_fill_brewer(palette = "Set3", name = "Age Group") +
  geom_text(
    position = position_dodge(width = .9),
    mapping = aes(label = (n), group = age_group_6cat),
    color = "#3b3b3b",
    vjust = -.5
  ) +
  labs(
    title = str_wrap(age_group_by_year_title,
      width = 120
    ),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths"
  ) +
  ylim(0, NA) +
  scale_x_discrete(breaks = analysis_year_range) +
  theme_suicide_data_report
```

## Death by Suicide and Race for Coconino County Residents (`r years_of_analysis`)

```{r 12-race-by-year}
# preparation
# color palette
# set color palettes
coconino_population_by_race_palette <- death_by_suicide_coconino_resident_race_specific_rate$race_code_lump

death_by_suicide <- azdhs_suicide_data_extract_plus_categorical_lumps

palette_race_4 <- brewer.pal(n = 4, name = "Set2")

names(palette_race_4) <- levels(factor(c(
  levels(death_by_suicide$race_code_lump),
  levels(coconino_population_by_race_palette)
))) # Extract all levels of both data

race_by_year_title <- str_c("Death by Suicide and Race (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
death_by_suicide_coconino_resident_race_specific_rate %>%
  filter(
    death_book_year %in% analysis_year_range
  ) %>%
  ggplot(mapping = aes(
    x = death_book_year,
    y = n,
    group = race_code_lump,
    color = race_code_lump
  )) +
  geom_line(size = 1) +
  geom_label(
    mapping = aes(
      fill = race_code_lump,
      x = death_book_year,
      y = n,
      label = (n)
    ),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  scale_color_manual(values = palette_race_4) +
  scale_fill_manual(values = palette_race_4) +
  ylim(0, NA) +
  labs(
    title = str_wrap(race_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    color = "Race",
    fill = "Race"
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Coconino County Resident Death by Suicide Rate by Race of Decedent (`r years_of_analysis`)

```{r 13-race-rate}
death_by_suicide_coconino_resident_race_specific_rate %>%
  filter(
    death_book_year %in% analysis_year_range,
    n > 6
  ) %>%
  ggplot(mapping = aes(
    x = death_book_year,
    y = rate_per_100k,
    group = race_code_lump,
    color = race_code_lump
  )) +
  geom_line(size = 1) +
  geom_label(
    mapping = aes(
      fill = race_code_lump,
      x = death_book_year,
      y = rate_per_100k,
      label = (rate_per_100k)
    ),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  scale_color_manual(values = palette_race_4) +
  scale_fill_manual(values = palette_race_4) +
  ylim(0,NA) +
  labs(
    title = str_wrap("Rate of Death by Suicide and Race (2017-2021)", width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Rate per 100,000*",
    color = "Race",
    fill = "Race",
    caption = "*Crude rate is not adjusted by age.
    Valid rates cannot be calculated when the count is less than 6."
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Percentage of Death by Suicide by Race of Decedent (`r years_of_analysis`)


```{r 14-race-by-percent}
# preparation
# Percent stacked barchart
death_by_suicide_race_percentage <- azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  count(race_code_lump) %>%
  mutate(n_percent = round(100 * (n / sum(n)), digits = 0)) %>%
  mutate(group = "Coconino Death by Suicide") %>%
  mutate(race_code = race_code_lump)

coconino_population_by_race_percentage <- azdhs_pop_data_by_sex_gender %>%
  filter(race_ethnicity != "All groups") %>%
  filter(year == 2020) %>%
  filter(age_group == "total") %>%
  filter(sex == "Total") %>%
  filter(area == "Coconino") %>%
  mutate(
    estimate = as.numeric(estimate),
    race_code_lump =
      case_when(
        race_ethnicity == "White non-Hispanic" ~ "White Non-Hispanic",
        race_ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino (any race)",
        race_ethnicity == "Black or African American" ~ "Other",
        race_ethnicity == "Asian or Pacific Islander" ~ "Other",
        race_ethnicity == "American Indian or Alaska Native" ~ "American Indian and Alaska Native"
      )
  ) %>%
  group_by(race_code_lump) %>%
  summarise(estimate = sum(estimate)) %>%
  mutate(
    n_percent = round(100 * (estimate / sum(estimate)), digits = 0),
    group = "Coconino 2020 Population"
  )

suicide_by_race_percentage <- bind_rows(
  death_by_suicide_race_percentage,
  coconino_population_by_race_percentage
)

suicide_by_race_percentage <- full_join(
  x = death_by_suicide_race_percentage,
  y = coconino_population_by_race_percentage,
  by = c(
    "n" = "estimate",
    "race_code_lump" = "race_code_lump",
    "n_percent" = "n_percent",
    "group" = "group"
  )
)

percent_race_by_year_title <- str_c("Percentage of Death by Suicide by Race in Comparison to 2020 Population (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
suicide_by_race_percentage %>%
  ggplot(mapping = aes(fill = race_code_lump, y = n_percent, x = group)) +
  geom_col(position = "fill") +
  geom_text(
    position = position_fill(vjust = 0.5),
    mapping = aes(label = comma(n)),
    color = "white"
  ) +
  scale_fill_manual(values = palette_race_4) +
  # scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = percent) +
  labs(
    title = str_wrap(percent_race_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "",
    y = "Percentage",
    fill = "Race",
    caption = str_wrap("The values presented within the plot represent the Coconino 2020 Population estimate and the number of deaths by suicide, respectively.", width = 120)
  ) +
  theme_suicide_data_report
```

## Deaths by Suicide and Gender for Coconino County Residents (`r years_of_analysis`)

```{r 15-sex-by-year}
# preparation
gender_by_year_title <- str_c("Deaths by Suicide by Sex of Decedent (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(sex) %>%
  ggplot(mapping = aes(x = death_book_year, y = n, group = sex, color = sex)) +
  geom_line(size = 1) +
  geom_label(
    mapping = aes(fill = sex, x = death_book_year, y = n, label = (n)),
    alpha = (5 / 6),
    color = "white",
    fontface = "bold"
  ) +
  ylim(0, NA) +
  scale_color_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  labs(
    title = str_wrap(gender_by_year_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    color = "Sex",
    fill = "Sex"
  ) +
  theme_suicide_data_report +
  guides(
    fill = guide_legend(
      override.aes = aes(label = "")
    )
  )
```

## Total Deaths by Suicide and Gender for Coconino County Residents (`r years_of_analysis`)

```{r 16-sex}
# preparation
gender_title <- str_c("Death by Suicide for Each Sex (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count() %>%
  ungroup() %>%
  ggplot(mapping = aes(x = sex, y = n)) +
  geom_col(mapping = aes(fill = sex), show.legend = FALSE) +
  geom_text(
    mapping = aes(label = n, y = n),
    color = "#525252",
    nudge_y = 8
  ) +
  scale_fill_manual(values = c("#7FC97F", "#BEAED4")) +
  ylim(0,NA) +
  labs(
    title = str_wrap(gender_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    fill = "Sex"
  ) +
  theme_suicide_data_report
```

## Means of Suicides in Coconino County (`r years_of_analysis`)

```{r 17-method}
# preparation
method_title <- str_c("Methods of Suicides (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(death_book_year) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  ggplot(mapping = aes(x = death_book_year, y = n, fill = method_code_cdc_lump)) +
  geom_bar(position = "dodge", stat = "identity") +
  ylim(0, NA) +
  geom_text(
    position = position_dodge(width = 1),
    mapping = aes(label = (n)),
    color = "#3b3b3b",
    vjust = -1
  ) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
    title = str_wrap(method_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Year",
    y = "Number of deaths",
    fill = "Method",
    caption = str_wrap(
      "Method is coded from a free text variable and is an injury description.",
      width = 120
    )
  ) +
  theme_suicide_data_report
```

## Means of Suicide by Gender in Coconino County (`r years_of_analysis`)

```{r 18-method-by-sex}
# preparation
method_by_sex_title <- str_c("Methods of Suicides by Sex of Decedent (",
  years_of_analysis,
  ")",
  sep = ""
)

# plot
azdhs_suicide_data_extract_plus_categorical_lumps %>%
  filter(
    death_book_year %in% analysis_year_range,
    county_resident == "Resident"
  ) %>% # county residents only %>%
  group_by(sex) %>%
  count(method_code_cdc_lump) %>%
  ungroup() %>%
  ggplot(mapping = aes(x = sex, y = n, group = method_code_cdc_lump, fill = method_code_cdc_lump)) +
  geom_bar(position = "dodge", stat = "identity") +
  ylim(0, NA) +
  geom_text(
    position = position_dodge(width = 1),
    mapping = aes(label = (n)),
    color = "#3b3b3b",
    vjust = -1
  ) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
    title = str_wrap(method_by_sex_title, width = 120),
    subtitle = "Including only Coconino County residents.",
    x = "Sex",
    y = "Number of deaths",
    fill = "Method",
    caption = str_wrap(
      "Method is coded from a free text variable and is an injury description.",
      width = 120
    )
  ) +
  theme_suicide_data_report
```

## Syndromic Surveillance (`r analysis_year_current`)

```{r 19-nssp}
# Profile info from parameters above in YAML header
myProfile <- Credentials$new(
  username = params$username,
  password = params$password
)

# # including this for testing and troubleshooting
# myProfile <- Credentials$new(
#   username = askme("Enter your username: "),
#   password = askme()
# )

## JSON URL from ESSENCE API
url <- ""

# Query definition:
# geography system: patient region
# patient region: coconino
# facility: banner page, FMC, Tuba City
# Medical Grouping: ESSENCE syndromes
# Discharge Dx Available: Yes
# CC and DD category: CDC suicidal ideation v1, CDC suicide attempt v1
# has been emergency: Yes
# admission type category: Emergency
# within graph stratification: CC and DD category
# graph display style: single graph
# remove zero series: true
# year stratification graph start week: Jan

## Get Data from ESSENCE
api_data <- get_api_data(url) # or api_data <- myProfile$get_api_data(url)

# tidy
essence_data <- api_data$timeSeriesData

# plot
essence_data %>%
  mutate(
    date = as.Date(date)
  ) %>%
  filter(date <= as.Date(Sys.Date()-7)) %>%
  ggplot(mapping = aes(
    x = date,
    y = count,
    color = lineLabel,
    group = lineLabel
  )) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b"
  ) +
  scale_color_manual(values = c("#C658E8", "#5886E8")) +
  ylim(0,NA) +
  labs(
    title = str_c("Number of Weekly Suicide Related ED Visits in ", analysis_year_current),
    subtitle = " Among Coconino County Residents",
    x = "Date",
    y = "Count",
    color = "CC DD Category",
    caption = str_wrap(
      string = "Each point represents one week. Data source: CDC NSSP BioSense Platform reporting data from - where a discharge diagnosis is available and it is an emergency admission.",
      width = 80
    )
  ) +
  theme_suicide_data_report
```

## About Syndromic Surveillance


**Syndromic Surveillance** is surveillance for *syndromes* as opposed to confirmed cases. **Syndromes** are defined by a specific set of concurrent *symptoms*. **Symptoms** are characteristics signs or indications of the existence of something else. In this case (see previous slide), emergency department visits related to suicide are included in the count.
